---
title: "p8105_HW2_jt3098"
author: "Jieqi Tu (jt3098)"
date: "9/29/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Problem 1
```{r importing data}
subway_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada)

subway_data$entry = ifelse(subway_data$entry == "YES", TRUE, FALSE)

```

This dataset is called subway_data. The variable it contains are below:

* line: the avenue or road in NYC that the station located in
* station_name: the names of all stations
* station_latitude: the latitude value for this station
* station_longitude: the longitude value for this station
* route1:route11: the train that stops at this station
* entry: whether or not there is entry for this station
* vending: whether or not there is vending for this station
* entrance_type: the type of the entrance
* ada: is there ada compliance in this station

Cleaning steps:

* Import csv file from the data folder
* Use janitor::clean_names() function to convert all upper-case letters to lower-case.
* Use select function to carefully choose columns required to show up in the dataset
* Use ifelse() function to convert entry variable from character to logical.

```{r distinctstations}
distinction = distinct(subway_data, station_name, line)
```


